Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-10 ICPP-19/communications.tex
Edit after Lionel's proofread

	\author{Valerie B\'eranger}
	\affiliation{%
		\institution{Inria Paris-Rocquencourt}
		\city{Rocquencourt}
		\country{France}
	}
	\author{John Smith}
	\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
	\email{jsmith@affiliation.org}

	\author{Julius P. Kumquat}
	\affiliation{\institution{The Kumquat Consortium}}
	\email{jpkumquat@consortium.net}

@@

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with finite memory capacity obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses, parallel resources, it is hard to obtain a precise model based on prediction of computation and communication timings. Fourth, number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tune optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form INRIA Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAPPI~\cite{kaapi} from INRIA Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information such as distance to exit (last) node as a priority or affinity of computation towards resources is also provided along with the DAG. Then runtime is responsible for managing scheduling of computations and communications, data transfers among different memories, computation-communication overlap, and load balance.

@@

	In the last few decades, there is a drastic improvement in the hardware to provide perminent rate of computation, but little improvement has been achieved for the rate of data movement. With extreme scale computing, supercomputers face bottlenecks due to the need of large amount of data~\cite{ascaccommitteereport2014,yelick2016}. Therefore focus of HPC community is now changing towards avoiding, hiding and minimizing communication costs.

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance, tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this paper. We prove that order of communications on two memory nodes with the objective of minimizing the makespan is NP-Complete. Our proof is inspired from work by ~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem. The main difference with their and our approach is that they consider a bounded number of tasks can await execution on second machine and buffer is acquired after completing the task on first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and memory is acquired before starting the data transfer on communication resource. We propose different strategies for runtimes such that it can maximize the overlap of computations and communications. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

	\renewcommand{\labelitemi}{$\bullet$}
	\begin{itemize}
		\item Proposed different scheduling strategies with the objective to minimize the makespan
		%%		\item An optimal strategy and its proof, when there is not any memory capacity constraint on the target machine
		\item NP-Completeness proof for the general data-transfer problem 
		\item Linear programming formulation of the problem
		\item Numerous experiments to assess the effectiveness of our strategies on molecular chemistry kernels 
	\end{itemize}

@@

	The problem of scheduling tasks has been highly studied in literature and this problem is known to be NP-Complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for this problem~\cite{heft-Topcuoglu,hipc16multiresource,ipdps16starpu}. There ia also a number of studies in the direction of task scheduling with the emphasis of improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al proposed a heuristic to schedule tasks on a computing resource where most of its data is available~\cite{luka-dmdar}. Similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et.al proposed heuristics to partition the task graph across a number of processors such that inter processor communication can be minimized ~\cite{predari:tel-01518956}.

	The problem considered in this paper also can be viewed as a flow shop problem. Communication and computation timings of a task can be considered as processing timings on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}. They consider a bounded number of tasks can await execution on second machine.

	There is also a numerous study pertaining to the scheduling with limited memory and storage, since the work of register allocation for arithmetic expressions by Sethi and Ulman~\cite{Sethi:1970:GOC:321607.321620}. Sarkar et. al worked on the scheduling of graphs of smaller-grain tasks with limited memory, where each task requires homogeneous data size ~\cite{vsarkar-pact}. The same work has been extend by Marchal et. al for task graphs where memory requirement of each task is highly irregular~\cite{loris-ipdps18}.

@@

	To exploit the full potential of a system tasks may get scheduled on processing
	units where all of their data does not reside. A task may require all of its data
	in local memory before starting the execution. There may be multiple tasks
	scheduled on a processing unit, which require to transfer data from the same
	memory node. Order of data transfers for such tasks is very crucial for the
	communication-computation overlap, thus for the overall performance. In general, order of task execution with input and output data transfers can be 
	viewed as 3-machine flowshop problem, where processing time on first machine is 
	input data transfer time, processing time on second machine is task execution time, 
	processing time on third machine is output data transfer time and the objective is 
	to minimize the total makespan. This is a well known hard problem~\cite{NPComplete3Machine,johnson}.

	In general, output data produced by a task is very less. It is often the case that future tasks running on the same memory node require output data of the past tasks. Therefore, most runtime systems transfer data to other memory nodes based on the demand -- not immediately. It is possible that all output data can be stored in a preallocated separate buffer on a memory node. Hence, we do not consider output data separately in our analysis and assume that output data is negligible or stored in a separate buffer for each task. Thus problem considered here is simpler than general 3-machine flowshop problem. We prove that order of execution of such tasks with finite memory capacity is an NP-complete problem. We also present an optimal algorithm to determine the order of execution of tasks when there is not any restriction on memory limit.

@@

	Let \scomm($i$) and  \scomp($i$) represent the start times of task $i$ on communication and computation resources. We call a task $i$ is compute (resp. communication) intensive if $CP_i $$\ge$ (resp. $<$) $CM_i$ ).

@@

	The algorithm presented in Section~\ref{sec:theoreticalProof} achieves optimal makespan when there is not any memory capacity constraint. This optimal value indicates a lower bound on the makespan. We use $optimal\_ \allowbreak makespan\_  \allowbreak infinite\_  \allowbreak memory$ ($OMIM$) to represent this value. We assess the efficiency of other proposed heuristics in limited memory capacity in Section~\ref{sec:exp} with respect to this lower bound, $OMIM$.

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. While in the second category, the best suited next task is chosen based on different criteria. In the last category, which is based on combination of strategies of the other two categories, we precompute the order and try to follow it as much as possible, but when the precomputed schedule induces idle time then we look for the best suited next task. In all of our strategies (except linear programming based strategy) order of communications and computations is same.

	\subsection{Static Ordering}
	In this class of strategies, we compute the order of processing in advance based on criteria such as communication time, compute time and accelerated tasks. After computing the order, we follow the same order on computation and communication resources and make sure that at each point in the schedule memory capacity is respected.

@@

	\begin{enumerate}[label=\roman*)]
		\item $order  \_ \allowbreak  \allowbreak of  \_ \allowbreak  \allowbreak optimal \_ \allowbreak strategy \_ \allowbreak infinite \_ \allowbreak memory $ ($OOSIM$): Order of processing is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. At each point in the schedule memory capacity is respected. Hence the makespan of this heuristic may be completely different from $OMIM$.

		\item $increasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak strategy$ ($IOCMS$): Order is obtained by sorting all tasks in non-decreasing order of communication time.

		\item $decreasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak computation \_ \allowbreak strategy$ ($DOCPS$): Order is obtained by sorting all tasks in non-increasing order of computation time. 
		\item $increasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak plus \_ \allowbreak computation \_ \linebreak[4] \allowbreak strategy$ ($IOCCS$): Order is obtained by sorting all tasks in non-decreasing order of sum of communication and computation times.
		\item $decreasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak plus \_ \allowbreak computation \_ \linebreak[4] \allowbreak strategy$ ($DOCCS$): Order is obtained by sorting all tasks in non-increasing order of sum of communication and computation times.

@@

	Fig~\ref{fig:staticOrderExample} shows schedules for all proposed heuristics of this class for the task set of Table~\ref{tab:staticOrderExample}.
	\subsection{Dynamic Selection}
	%%%%string alg_name="largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";

	In this class of strategy, when communication resource is idle, a task is chosen which satisfies the selection criteria maximally among other tasks. For example, if the selection criteria is to choose a highly compute intensive task, then we calculate the ratio of compute time and communication time for all tasks, and the task with the maximum ratio is selected (ties are broken randomly). As we assume that order on both resources is same, hence order on compute resource follows the order of communication resource.

	We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		%%	\item $largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the largest communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: A task with the smallest communication time is chosen. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the maximum ratio of compute time to communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory.

		\item $largest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($LCMR$): A task with the largest communication time is chosen. 
		\item $smallest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($SCMR$): A task with the smallest communication time is chosen.
		\item $maximum \_ \allowbreak accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ \linebreak[4]($MAMR$): A task with the maximum ratio of compute time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

@@

	Fig~\ref{fig:dynamicSelectionExample} shows schedules for all proposed heuristics of this class for the task set of Table~\ref{tab:dynamicSelectionExample}.

@@

	In this class of strategy, we precompute the order of tasks based on different criteria and then try to follow the same order as much as possible. But when communication resource is idle due to memory requirement of the next task in the obtained order, then we select a task similar to previous strategy. After a task is selected then we update the remaining order without this task. This class of strategy takes advantage of static information in the form of precomputed order and dynamic corrections to minimize the idle time due to memory constraint.
	%%%%alg_name="optimal_order_infinite_memory_largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";

	Initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. When communication resource is idle due to memory requirement of the next task in the order, then we select a set of tasks which induce minimum idle time on compute resource and satisfy memory requirement. We define the following heuristics based on how we select a task from this set. After that order of remaining tasks is updated without this task.

	\begin{enumerate}[label=\roman*)]
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak largest \_ \allowbreak communication \_ \linebreak[4]\allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOLCMR$): A task with the largest communication time is chosen from the set.
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak smallest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOSCMR$): A task with the smallest communication time is chosen from the set.
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak maximum \_ \allowbreak accelerated \_ \linebreak[4]\allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOMAMR$): A task, whose ratio of compute time to communication time is maximum, is chosen from the set .
	\end{enumerate}

@@

	Fig~\ref{fig:staticOrderDynamicCorrectionsExample} shows schedules for all proposed heuristics of this class for the task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample}.

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations.  $COMP_i$ and $COMM_i$ represent computation and communication timings of task $i$. Memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent start and end time of communication (resp. computation) for task $i$. The formulation contains $i)$ n-1 boolean variables, $a_{ij}$, for each task $i$ to denote the order of $i$ and $j$ on communication resource $ii)$ n-1 boolean variables, $b_{ij}$, for each task $T_i$ to denote the order of $i$ and $j$ on computation resource, and $iii)$ n-1 boolean variables, $c_{ij}$, for each task $i$ to denote the order of $s_i$ and $e'_j$.

@@

	\noindent Let $L=\sum_i (COMP_i + COMM_i)$. It is evident that $e_i =s_i + COMM_i$ and $e'_i =s'_i + COMP_i$. The constraints are following.

>>>>

	\author{Suraj Kumar}
	\affiliation{%
		\institution{Pacific Northwest National Laboratory}
		\city{Richland, Washington, USA}}
	\email{suraj.kumar@pnnl.gov}

	\author{Lionel Eyraud-Dubois}
	\affiliation{\institution{Inria Bordeaux -- Sud-Ouest}\institution{Université de Bordeaux, France}}
	\email{lionel.eyraud-dubois@inria.fr}

	\author{Sriram Krishnamoorthy}
	\affiliation{
	\institution{Pacific Northwest National Laboratory}
	\city{Richland, Washington, USA}}
	\email{sriram@pnnl.gov}

@@

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with finite memory capacity, obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges that come into picture. First, every architecture provides its own efficiency and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses and parallel resources, it is challenging to obtain a precise model based on prediction of computation and communication timings. Fourth, then number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tuned optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form Inria Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAAPI~\cite{kaapi} from Inria Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information is also provided along with the DAG, such as distance to exit (last) node as a priority or affinity of computation towards resources. The runtime is then responsible for managing scheduling of computations and communications, data transfers among different memories, computation-communication overlap, and load balance.

@@

	In the last few decades, we have witnessed a drastic improvement in the hardware to provide higher rate of computation, but comparatively smaller improvement has been achieved for the rate of data movement. With extreme scale computing, supercomputers face bottlenecks due to the need of large amount of data~\cite{ascaccommitteereport2014,yelick2016}. Therefore, the HPC community is now focusing on avoiding, hiding and minimizing communication costs.

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of independent ready tasks and schedules them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computations and communications. This is the main topic of this paper. We prove that the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-Complete problem if the memory of the target node is limited. Our proof is inspired from work by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider that all tasks have the same occupation on the second machine) and that the memory occupation starts when the second machine starts working. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on communication resource. We propose different runtime strategies in order to maximize the overlap of computations and communications. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper:

	\renewcommand{\labelitemi}{$\bullet$}
	\begin{itemize}
		\item NP-Completeness proof for the general data-transfer problem 
		\item Proposed different scheduling strategies with the objective to minimize the makespan
		%%		\item An optimal strategy and its proof, when there is not any memory capacity constraint on the target machine
		\item Linear programming formulation of the problem
		\item Numerous experiments to assess the effectiveness of our strategies on molecular chemistry kernels 
	\end{itemize}

@@

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-Complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,hipc16multiresource,ipdps16starpu}. There is also a number of studies in the direction of task scheduling with the emphasis of improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computing resource where most of its data is available. A similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et. al proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

	The problem considered in this paper also can be viewed as a flow shop problem: the communication link can be seen as a processing resource, and each task needs to first be handled by the communication link and then by the computational resource. Communication and computation timings of a task can thus be considered as processing times on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}, in which a constraint is imposed on the number of tasks that can await execution on the second machine.

	A number of other studies have focused on scheduling with limited memory and storage, starting with the work of register allocation for arithmetic expressions by Sethi and Ulman~\cite{Sethi:1970:GOC:321607.321620}. Sarkar et. al worked on the scheduling of graphs of smaller-grain tasks with limited memory, where each task requires homogeneous data size~\cite{vsarkar-pact}. The same work has been extend by Marchal et. al for task graphs where memory requirement of each task is highly irregular~\cite{loris-ipdps18}.

@@

	To exploit the full potential of a system it may be necessary to execute tasks on processing
	units where all of their data does not reside. A task may require all of its input data
	in local memory before starting the execution. There may be multiple tasks
	scheduled on a processing unit, which require to transfer data from the same
	memory node. Ordering data transfers for such tasks is very crucial for the
	communication-computation overlap, thus for the overall
	performance. In general, order of task execution with input
	and output data transfers can be  
	viewed as a 3-machine flowshop problem, where processing time on the first machine is 
	input data transfer time, processing time on the second machine is task execution time, 
	and processing time on the third machine is output data transfer time; and the objective is 
	to minimize the total makespan. This is a well known hard problem~\cite{NPComplete3Machine,johnson}.

	In many cases, output data that needs to be retrieved after task execution is much smaller than the input data. It is often the case that future tasks running on the same memory node require output data of the past tasks. Therefore, most runtime systems transfer data to other memory nodes based on the demand -- not immediately after they were produced. It is also possible that all output data can be stored in a preallocated separate buffer on a memory node. Hence, we do not consider output data explicitely in our analysis and assume that output data is negligible or stored in a separate buffer for each task. Thus problem considered here is thus more similar to a 2-machine flowshop problem. We prove that ordering the execution of such tasks with finite memory capacity is an NP-complete problem: %%We also present an optimal algorithm to determine the order of execution of tasks when there is not any restriction on memory limit.

@@

	Given a schedule, \scomm($i$) and \scomp($i$) represent the
	start times of task $i$ on communication and computation
	resources. A schedule is feasible if for every time $t$, the
	amount of memory required by tasks such that $\scomm(i) \leq t
	\leq \scomp(i) + COMP_i$ is not more than the memory capacity
	$C$. For simplicity, we assume in all of this paper that the
	amount of memory required by a task is proportional to its
	communication time, and without loss of generality we consider
	that memory size of a task is equal to its communication time.

	We call a task $i$ compute intensive if $COMP_i \ge CM_i$, and
	communication intensive otherwise.

@@

	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} presented in Section~\ref{sec:theoreticalProof} achieves an optimal makespan when there is no memory constraint. This optimal value indicates a lower bound on the makespan of the constrained case. We use denote this value with $optimal\_ \allowbreak makespan\_  \allowbreak infinite\_  \allowbreak memory$ ($OMIM$). In the present Section, we propose different heuristics for the limited memory case, and we assess their efficiency with respect to this lower bound in Section~\ref{sec:exp}.

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dyanmically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static ordering is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computations take place in the same order.

	\subsection{Static Ordering}
	In this class of strategies, we compute the order of processing in advance based on criteria such as communication time, compute time and accelerated tasks. After computing the order, we follow the same order on computation and communication resources and make sure that the memory constraint is respected at each point in the schedule.

@@

	\begin{enumerate}[label=\roman*)]
		\item \textit{Order  of  optimal strategy infinite memory} ($OOSIM$): This heuristic uses the order given by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, but respects the memory constraint at each point in the schedule. Hence the makespan of this heuristic may be completely different from $OMIM$.

		\item \textit{increasing order of communication strategy} ($IOCMS$): tasks are ordered  in non-decreasing order of communication time.

		\item \textit{decreasing order of computation strategy} ($DOCPS$): tasks are ordered in non-increasing order of computation time. 
		\item \textit{increasing order of communication plus computation strategy} ($IOCCS$): tasks are ordered in non-decreasing order of the sum of their communication and computation times.
		\item \textit{decreasing order of communication plus computation strategy} ($DOCCS$): tasks are ordered in non-increasing order of the sum of their communication and computation times.

@@

	In order to highlight the different behavior of these static strategies, we propose on Table~\ref{tab:staticOrderExample} an example instance, and on Figure~\ref{fig:staticOrderExample} the corresponding schedules for all these heuristics.

	\subsection{Dynamic Selection}
	%%%%string alg_name="largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";

	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a ranking criterion which differ depending on the heuristic, among those which fit in memory. For example, if the selection criteria is to choose a highly compute intensive task, then we compute the ratio of compute time and communication time for all tasks, and we select the task which fit in the currently available memory and whose ratio is maximum (ties are broken arbitrarily). We also ensure that the order on both resources is the same, by ordering tasks on the computing resource accordingly.

	\todo[inline]{LED: I added this part ``among those which fit in memory'', is it correct?}
	\todo[inline]{LED: What happens if no task fit in memory?}

	%% We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		%%	\item $largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the largest communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: A task with the smallest communication time is chosen. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the maximum ratio of compute time to communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory.

		\item \textit{largest communication task respects memory restriction} ($LCMR$): a task with the largest communication time is chosen. 
		\item \textit{smallest communication task respects memory restriction} ($SCMR$): a task with the smallest communication time is chosen.
		\item \textit{maximum accelerated task respects memory restriction} ($MAMR$): a task with the maximum ratio of compute time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

@@

	We highlight the different dynamic heuristics with the instance described on Table~\ref{tab:dynamicSelectionExample} (these heuristics are too similar on the instance from the previous class), and Figure~\ref{fig:dynamicSelectionExample} shows the corresponding schedules.

@@

	In this class of strategy, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is dynamically selected, we update the remaining order without this task. This class of strategy takes advantage of static information in the form of precomputed order and dynamic corrections to minimize the idle time due to memory constraint.
	%%%%alg_name="optimal_order_infinite_memory_largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";

	In all strategies, the initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. . We define the following heuristics based on how we select a task from the set of tasks which fit in memory.
	\todo[inline]{LED: What happens if no task fit in memory?}

	\begin{enumerate}[label=\roman*)]
		\item \textit{optimal order infinite memory largest communication task respects memory restriction} ($OOLCMR$): a task with the largest communication time is chosen from the set.
		\item \textit{optimal order infinite memory smallest communication task respects memory restriction} ($OOSCMR$): a task with the smallest communication time is chosen from the set.
		\item \textit{optimal order infinite memory maximum accelerated task respects memory restriction} ($OOMAMR$): a task  with the highest ratio of compute time to communication time is chosen from the set.
	\end{enumerate}

@@

	As previously, we propose on Table~\ref{tab:staticOrderDynamicCorrectionsExample} an example instance for this class of strategies, and provide on Figure~\ref{fig:staticOrderDynamicCorrectionsExample} the corresponding schedules.

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations. Recall that $COMP_i$ and $COMM_i$ represent computation and communication timings of task $i$, and that the memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent the start and end times of communication (resp. computation) for task $i$. The formulation also contains for each pair of tasks $i$ and $j$ \textit{i)} a boolean variable $a_{ij}$ to denote the order of $i$ and $j$ on the communication resource \textit{ii)} a boolean variables $b_{ij}$ to denote the order of $i$ and $j$ on the computation resource, and \textit{iii)} a boolean variables $c_{ij}$ to denote the order of $s_i$ and $e'_j$.

@@

	\noindent Let $L=\sum_i (COMP_i + COMM_i)$. It is evident that $e_i =s_i + COMM_i$ and $e'_i =s'_i + COMP_i$. The linear program is given below.

	\vspace*{-0.5cm}
	\begin{align*}
		& \text{Minimize $l$ subject to:}\\
		\forall i, \quad & e'_i \leq l && \text{(task $i$ completes)}\\
		\forall i, \quad & e_i \leq s'_i&& \text{(valid ordering for task $i$)}\\
		\forall i, \forall j\ne i, \quad & \left\{\begin{aligned}
			e_j & \leq s_i +(1-a_{ij})L\\
			e_i & \leq s_j +a_{ij}L 
		\end{aligned} \right. &&
		\begin{aligned}
			&\text{(exclusive use of}\\
			&\text{  communication link)}
		\end{aligned}\\
		\forall i, \forall j\ne i, \quad & \left\{
		\begin{aligned}
			e'_j &\leq s'_i +(1-b_{ij})L\\
			e'_i &\leq s'_j +b_{ij}L
		\end{aligned}\right. && 
		\begin{aligned}
			&\text{(exclusive use of}\\
			&\text{  computation resource)}
		\end{aligned}\\
		\forall i, \forall j\ne i, \quad &\left\{
		\begin{aligned}
			e'_j &\leq s_i +(1-c_{ij})L\\
			s_i  &< e'_j +c_{ij}L\\
		\end{aligned}\right.&&
		\begin{aligned}
			&\text{(respect ordering}\\
			&\text{ of variables $c_{ij}$)}
		\end{aligned}\\
		\forall i, \quad & \sum_{r\ne i} (a_{ir} - c_{ir})CM(r) + CM(i) \le C && \text{(memory constraint)}\\
		%% %%\forall i, \forall j\ne i, \quad & \text{task }T_i \text{ respects memory capacity on computation resource:}\\
		%% %%& s_j < e'_i +(1-d_{ij})L\\
		%% %%& e'_i\leq s_j +d_{ij}L\\
		%% %%& \sum_r (d_{ir} - b_{ir})CM(r) + TM(i) \le C\\
	\end{align*}
